---
title: "Hw2"
author: "Areeya Aksornpan, Zayd Abdalla"
date: "3/15/2021"
output: pdf_document
---

## R Markdown


```{r}
library(mosaic)
library(mosaicData)
library(tidyverse)
library(tidymodels)
library(tidyquant)
library(estimatr)
library(gcookbook)
library(ggthemes)
library(kknn)
library(glmnet)
library(lubridate)
library(scales)
library(patchwork)
library(hrbrthemes)
library(kableExtra)
```
## Problem 1: visualization
```{r}

read_data <- function(df) {
  path <- paste("https://raw.githubusercontent.com/jgscott/ECO395M/master/data/", 
                     df, sep = "")
  df <- read_csv(path)
  return(df)
}

capmetro <- read_data("capmetro_UT.csv") %>%
  mutate(day_of_week = factor(day_of_week,
                              levels = c("Mon", "Tue", "Wed","Thu",
                                         "Fri","Sat","Sun")),
         month = factor(month, levels = c("Sep", "Oct","Nov")))

Figure1 <- 
  capmetro %>%
  group_by(hour_of_day, day_of_week, month) %>%
  mutate(avg_boarding = mean(boarding)) %>%
  ungroup() %>%
  ggplot() +
  geom_line(aes(x = hour_of_day, y = avg_boarding, color = month)) +
  scale_x_continuous(expand = c(0,0), limits = c(0, 24), 
                     breaks = seq(10, 20, 5)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, 200)) +
  scale_color_ft("Month") +
  facet_wrap(. ~ day_of_week, scales = "free") + 
  labs(x = "Hour of day", y = "Average boarding",
       title = "Average bus ridership around UT",
       subtitle = "Tracked by Optical Scanner",
       caption = "Source: Capital Metro") + 
  theme_ipsum(grid = "XY", axis = "xy") 

Figure1
```
#Figure1 illustrates the average CapMetro bus boardings-tracked by Optical Scanner-on weekdays in September, October, and November. The hour of peak boarding appears broadly similar across days, generally peaking around the 17th hour (5pm). This result is intuitive since most people finish school/work around that time. However, weekends tend to not peak in average bus boardings around certain hours as sharply, which supports my intuition that these trends are indicating work commutes. One guess for the decline in average boardings on Mondays in September is that the first Monday of September is Labor Day. Since Labor Day is a holiday, work commutes that day will decline relative to other Mondays, so the average bus boardings in September declines. One guess for the decline in average boardings on Weds/Thurs/Fri in November are because many schools and occupations go on break after the Tuesday before Thanksgiving, which gives people time off from work, so they are less likely to commute on those days.

```{r}
Figure2 <-
  capmetro %>%
  group_by(timestamp, hour_of_day) %>%
  mutate(avg_boarding = mean(boarding)) %>%
  ggplot() +
  geom_point(aes(x = temperature, y = avg_boarding, color = weekend)) +
  scale_x_continuous(expand = c(0,0), limits = c(30, 100), 
                     breaks = seq(40, 100, 20)) +
  scale_y_continuous(expand = c(0,0), limits = c(0, 300)) +
  scale_color_ft() +
  facet_wrap(. ~ hour_of_day, scales = "free") +  
  labs(x = "Temperature", y = "Boarding",
       title = "Average bus ridership around UT by temperature",
       subtitle = "Faceted by hour of day",
       caption = "Source: Capital Metro") + 
  theme_ipsum(grid = "XY", axis = "xy") +
  theme(legend.title = element_blank()) 

Figure2
```
#Figure2 shows average ridership, by temperature, which is faceted by hour of the day (6am to 10pm), and averaged by 15-minute increments. Gray indicates weekdays and blue indicates weekends. When we hold hour of day and weekend status constant, temperature does not appear to noticeably change the average ridership of UT students. The changes in bus demand seems to be more related to the time of day since the average boardings at each hour is pretty similar across temperatures. 

## Problem 2: Saratoga House Prices

```{r}

saratoga <- mosaicData::SaratogaHouses

#create the train/test split.

set.seed(300)

saratoga_split <- initial_split(saratoga, strata = "price", prop = 0.75)
saratoga_train <- training(saratoga_split)
saratoga_test  <- testing(saratoga_split)

dim(saratoga_train)
dim(saratoga_split)

#use cross-validation to split training set into k-folds.

# 3 fold cross validation 
saratoga_fold <- vfold_cv(saratoga_train, v = 3, repeats = 1, strata = "price")

# Linear and Knn models
lin_mod <-
    linear_reg() %>%
    set_mode("regression") %>%
    set_engine("lm")
lin_mod

knn_mod <-
  nearest_neighbor(
    mode = "regression",
    neighbors = tune("K"),
  ) %>%
  set_engine("kknn")
knn_mod

#Use tidymodels to feature engineer: rescaling and standardizing variables
saratoga_wf <-
  workflow() %>%
  add_formula(price ~ .) %>%
  # log price
  step_log(price) %>%
  # mean impute numeric variables
  step_meanimpute(all_numeric(), -all_outcomes()) %>%
  # rescale all numeric variables to lie between 0 and 1
  step_range(all_numeric(), min = 0, max = 1) %>%
  # one-hot
  step_dummy(fuel, centralAir, heating, newConstruction, waterfront, sewer) %>%
  # remove predictor variables that are almost the same for every entry
  step_nzv(all_predictors()) 
saratoga_wf

#Fitting LM model

set.seed(400)
lm_rs <- 
  saratoga_wf %>%
  add_model(lin_mod) %>%
  fit_resamples(
    resamples = saratoga_fold,
    control = control_resamples(save_pred = TRUE)
  )


#Fitting KNN model


set.seed(400)
# feature engineering
knn_recipe <- 
  recipe(price ~ ., data = saratoga_train) %>%
  # log price
  step_log(price) %>%
  # mean impute numeric variables
  step_meanimpute(all_numeric(), -all_outcomes()) %>%
  # rescale all numeric variables to lie between 0 and 1
  step_range(all_numeric(), min = 0, max = 1) %>%
  # one-hot
  step_dummy(fuel, centralAir, heating, newConstruction, waterfront, sewer) %>%
  # remove predictor variables that are almost the same for every entry
  step_nzv(all_predictors()) 
# workflow
knn_wf <- 
  workflow() %>%
  add_model(knn_mod) %>%
  add_recipe(knn_recipe)
# hyperparameter tuning
gridvals <- tibble(K = seq(1, 200))

knn_rs <- 
  knn_wf %>%
  tune_grid(
    resamples = saratoga_fold,
    grid = gridvals,
    control = control_resamples(save_pred = TRUE)) 
knn_rs

set.seed(400)

# Display only minimum RMSE
knn_min <- knn_rs %>%
  collect_metrics() %>% 
  filter(.metric == "rmse") %>%
  filter(mean == min(mean))
knn_min


## Evaluate Models

# Evaluate Linear Model
final_lm_wf <- 
  saratoga_wf %>%
  add_model(lin_mod) 
  
lm_fit <- 
  final_lm_wf %>%
  last_fit(split = saratoga_split)
lm_fit %>% collect_metrics()
lm_results <-
  lm_fit %>%
  collect_predictions()
# view results
lm_results


lm_fit$.workflow[[1]] %>% 
  tidy() %>% 
  kable(digits = 4, "pipe") 

# LM Graphically
lm_results %>%
  ggplot(aes(.pred, price)) +
  geom_abline(lty = 2, color = "black", size = 1) +
  geom_point(alpha = 0.5, color = "dark green") +
  labs(
    title = 'Linear Regression Results',
    x = "Truth",
    y = "Predicted price",
    color = NULL
  ) + 
  theme_ipsum()

# Evaluate KNN Model

final_knn_wf <- 
  knn_wf %>% 
  finalize_workflow(knn_min)
knn_fit <- 
  final_knn_wf %>% 
  last_fit(split = saratoga_split)
knn_fit %>% collect_metrics()

# predictions
knn_results <- 
  knn_fit %>% 
  collect_predictions()
# view results
knn_results

# KNN Graphically
knn_results %>%
  ggplot(aes(.pred, price)) +
  geom_abline(lty = 2, color = "black", size = 1) +
  geom_point(alpha = 0.5, color = "dark green") +
  labs(
    title = 'KNN Regression Results',
    x = "Truth",
    y = "Predicted price",
    color = NULL
  ) + 
  theme_ipsum()
```
# We built two models-a linear model and a KNN model-to predict the price of houses. The base model appeared to perform quite well, so we decided to tweak it by feature engineering to improve the accuracy. We standardized numeric variables to values between (0,1), applied a log transformation to the price variable and created dummy variables for all "character" encoded variables. Next, both our linear and KNN regression models were trained on 3 folds without repetition. We gave the KNN model a hyperparameter (neighbors) that was tuned using a tuning grid. Then, these models were fit on out-of-sample data and we found that our linear model clearly outperformed the medium model from class. However, our KNN model heavily outperformed even our improved linear model. This exercise illustrates the capability of KNN models to adapt to non-linearities of the data in order to achieve better fits in predicting pricing of houses. 

## Problem 3: Classification and retrospective sampling
```{r}

german_credit <- 
  read_data("german_credit.csv") %>% 
  select(-1) %>%
  # Factoring outcomes 
  mutate(Default = as.factor(Default))

# Build logistic regression model


german_credit %>%
  group_by(Default, history) %>%
  add_tally() %>% 
  rename(num_default = n) %>% 
  distinct(history, num_default) %>%
  ungroup() %>%
  group_by(history) %>%
  mutate(tot_default = sum(num_default),
         prob_default = (num_default / tot_default) * 100) %>%
  filter(Default == 0) %>%
  ggplot() +
  geom_col(aes(x = history, y = prob_default, 
             fill = history)) + 
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_x_discrete(labels = c("Good", "Poor", "Terrible")) +
  scale_fill_ft() +
  labs(x = "History", y = "Probability of Default",
       title = "Probability of credit default by credit history") +
  theme_ipsum(grid = "Y") + 
  theme(legend.title = element_blank(),
        legend.position = "None")




# Train test
set.seed(395)
german_split <- initial_split(german_credit, strata = "Default", prop = 0.75)
german_train <- training(german_split)
german_test  <- testing(german_split)
# 3 fold cross validation (for speed)
german_fold <- vfold_cv(german_train, v = 3, repeats = 1, strata = "Default")
german_fold 


# Model engine

log_mod <-
  logistic_reg(penalty = tune(), mixture = 1) %>% 
  set_engine("glmnet") %>% 
  set_mode("classification") 
log_mod


# recipe and workflow.

set.seed(350)
# varlist to keep
varlist <- c("Default", "duration", "amount", "installment", "age", 
             "history", "purpose", "foreign")
# recipe
log_rec <- 
  recipe(Default ~ ., data = german_train) %>%
  # remove vars not in varlist
  step_rm(setdiff(colnames(german_credit), varlist)) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors())
  
# workflow
log_wf <- 
  workflow() %>%
  add_model(log_mod) %>%
  add_recipe(log_rec)


# Tune grid


log_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))

set.seed(350)
log_rs <- 
  log_wf %>% 
  tune_grid(german_fold,
            grid = log_grid,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(roc_auc))
log_rs

log_rs %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  ylab("Area under the ROC Curve") +
  scale_x_log10(labels = scales::label_number()) + 
  theme_ipsum()

top_models <-
  log_rs %>% 
  show_best("roc_auc", n = 20) %>% 
  arrange(penalty) 
top_models %>% kbl(format = "pipe", booktabs = T)

log_rs %>%
  select_best()

# Model 10 seems to be the best
# Graphically 
log_best <- 
  log_rs %>% 
  collect_metrics() %>% 
  arrange(penalty) %>% 
  slice(10)
log_auc <- 
  log_rs %>% 
  collect_predictions(parameters = log_best) %>% 
  roc_curve(Default, .pred_0) %>% 
  mutate(model = "Logistic Regression")
autoplot(log_auc)

final_log_wf <- 
  log_wf %>% 
  finalize_workflow(log_best)
log_fit <- 
  final_log_wf %>% 
  last_fit(split = german_split)
log_fit %>% collect_metrics()
log_results <- 
  log_fit %>% 
  collect_predictions()

log_results

log_results %>% 
  roc_curve(Default, .pred_0) %>% 
  autoplot()


# Confusion matrix
cm <- log_results %>%
  conf_mat(Default, .pred_class)
cm

# Poor sampling
german_credit %>% 
  group_by(history) %>% 
  tally() %>%
  kbl(format = "pipe")

```
# Our model is accurate roughly 74.4 percent of the time, which is not ideal since our null model that assumes no one will default would be correct 70 percent of the time. We believe the data is not likely ideal for predicting due to the weight of the poorly sampled history variable. Specifically, observe the vast disparity in sampling above.

## Problem 4: Children and Hotel Reservations
```{r}

hotels_dev <- 
  read_data("hotels_dev.csv") %>%
  mutate(children = as.factor(children))

  hotels_dev %>%
  count(children) %>%
  mutate(prop = round( n/sum(n), 3)) %>%
  mutate(children = if_else(children == 1, "children", "none")) %>%
  kbl("pipe")


# Children only make up about 8% of the sample

hotel_splits <- initial_split(hotels_dev, strata = children)
hotel_train <- training(hotel_splits)
hotel_test <- testing(hotel_splits)

train_val_set <- validation_split(hotel_train, strata = children, prop = 0.8)

# Proportion of children in train/test
# train
hotel_train %>%
  count(children) %>%
  mutate(prop = round( n/sum(n), 3)) %>%
  mutate(children = if_else(children == 1, "children", "none")) %>%
  kbl("pipe")

# test
hotel_test %>%
  count(children) %>%
  mutate(prop = round( n/sum(n), 3)) %>%
  mutate(children = if_else(children == 1, "children", "none")) %>%
  kbl("pipe")


# Both the splits are similar in proportion for children and no children.

## Baseline models

# penalized logistic regression model 

log_mod_base1 <- 
  logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

# Preprocess recipe
log_mod_base1_recipe <-
  recipe(children ~ market_segment + adults + customer_type + is_repeated_guest,
         data = hotel_train) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())

# Preprocess
log_mod_base1_wrkflow <-
  workflow() %>%
  add_model(log_mod_base1) %>%
  add_recipe(log_mod_base1_recipe)


# Tune hyperparameter

lr_reg_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))
log_base1_res <- 
  log_mod_base1_wrkflow %>%
  tune_grid(train_val_set,
            grid = lr_reg_grid,
            control = control_grid(save_pred = T),
            metrics = metric_set(roc_auc))


# select the best model
log_base1_res %>% 
  select_best()
best_mod_base1 <- 
  log_base1_res %>%
  collect_metrics() %>%
  slice(11)

# roc curve
log_base1_res %>%
  collect_predictions(parameters = best_mod_base1) %>%
  roc_curve(children, .pred_0) %>% 
  mutate(model = "Logistic Regression") %>%
  autoplot() 

# This is awful
# Confusion matrix

param_final <- 
  log_base1_res %>%
  select_best(metric = "roc_auc")
log_mod_base1_wrkflow <-
  log_mod_base1_wrkflow %>% 
  finalize_workflow(param_final)
base1_fit <- 
  log_mod_base1_wrkflow %>%
  last_fit(hotel_splits)
base1_pred <-
  base1_fit %>%
  collect_predictions()
base1_pred %>% 
  conf_mat(truth = children, estimate = .pred_class)


# Baseline 2

# Preprocess recipe
holidays <- c("AllSouls", "AshWednesday", "ChristmasEve", "Easter", 
              "ChristmasDay", "GoodFriday", "NewYearsDay", "PalmSunday")
log_mod_base2_recipe <-
  recipe(children ~ .,
         data = hotel_train) %>%
  
  step_date(arrival_date) %>% 
  step_holiday(arrival_date, holidays = holidays) %>% 
  step_rm(arrival_date) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>%
  
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())
# Preprocess
log_mod_base2_wrkflow <-
  workflow() %>%
  add_model(log_mod_base1) %>%
  add_recipe(log_mod_base2_recipe)


# Tune Hyperparameter

log_base2_res <- 
  log_mod_base2_wrkflow %>%
  tune_grid(train_val_set,
            grid = lr_reg_grid,
            control = control_grid(save_pred = T),
            metrics = metric_set(roc_auc))


log_base2_res %>% 
  collect_metrics() %>% 
  ggplot(aes(x = penalty, y = mean)) + 
  geom_point() + 
  geom_line() + 
  ylab("Area under the ROC Curve") +
  scale_x_log10(labels = scales::label_number()) + 
  theme_clean()


# select the best model
log_base2_res %>% 
  select_best()
best_mod_base2 <- 
  log_base1_res %>%
  collect_metrics() %>%
  slice(14)

# roc curve
log_base2_res %>%
  collect_predictions(parameters = best_mod_base2) %>%
  roc_curve(children, .pred_0) %>% 
  mutate(model = "Logistic Regression") %>%
  autoplot() 


# Not great, but better than baseline1 at least

param_final <- 
  log_base2_res %>%
  select_best(metric = "roc_auc")
log_mod_base2_wrkflow <-
  log_mod_base2_wrkflow %>% 
  finalize_workflow(param_final)
base2_fit <- 
  log_mod_base2_wrkflow %>%
  last_fit(hotel_splits)
base2_pred <-
  base2_fit %>%
  collect_predictions()
base2_pred %>% 
  conf_mat(truth = children, estimate = .pred_class)

# Better


## Best Linear Model


set.seed(400)
# train/test

hotel_splits2 <- initial_split(hotels_dev, strata = children)
hotel_train2 <- training(hotel_splits2)
hotel_test2 <- testing(hotel_splits2)

# cross-val folds
hotel_cv <- vfold_cv(hotel_train2, v = 10, repeats = 1, strata = children)

# validation set
hotels_val <- read_data("hotels_val.csv") %>%
  mutate(children = as.factor(children))
log_mod_rec <-
  recipe(children ~ .,
         data = hotel_train2) %>%
  
  step_date(arrival_date) %>% 
  step_holiday(arrival_date, holidays = timeDate::listHolidays("US")) %>% 
  step_rm(arrival_date) %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())
log_mod <- 
  logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

# Preprocess
log_mod_wrkflow <-
  workflow() %>%
  add_model(log_mod) %>%
  add_recipe(log_mod_rec)

# fit to  validation set
hotel_res <-
  log_mod_wrkflow %>%
  tune_grid(grid = lr_reg_grid,
            resamples = hotel_cv,
            control = control_grid(save_pred = T),
            metrics = metric_set(roc_auc))
top_models <-
  hotel_res %>% 
  show_best("roc_auc", n = 15) %>% 
  arrange(penalty) 
hotel_best <- 
  hotel_res %>% 
  collect_metrics() %>% 
  arrange(penalty) %>% 
  slice(12)

hotel_best %>% kbl("pipe")


# roc curve
hotel_res %>%
  collect_predictions(parameters = hotel_best) %>%
  roc_curve(children, .pred_0) %>% 
  mutate(model = "Logistic Regression") %>%
  autoplot() 


### Model Validation 1 using new dataset

trained_wf <-
  log_mod_wrkflow %>%
  finalize_workflow(hotel_best) %>%
  fit(hotels_dev)
hotel_preds <-
  trained_wf %>%
  predict(hotels_val) %>%
  bind_cols(hotels_val %>% select(children))
hotel_final_pred <- 
  trained_wf %>%
  predict(hotels_val, type = "prob") %>%
  bind_cols(hotel_preds)
hotel_final_pred %>%
  conf_mat(truth = children, .pred_class)
hotel_final_pred %>% 
  roc_curve(children, .pred_0) %>% 
  mutate(model = "Logistic Regression") %>%
  autoplot()



### Model Validation 2

# create v-folds
hotel_folds <- vfold_cv(hotels_val, v = 20)


hotel_fold_fit <- 
  trained_wf %>%
  fit_resamples(
    resamples = hotel_folds,
    control = control_resamples(save_pred = TRUE)
  )


# predicted probabilities
pred_sums <- list()
for (i in 1:20) {
  pred_sums <-
    hotel_fold_fit$.predictions[[i]] %>%
    summarize(sum_pred = sum(as.numeric(.pred_class))) %>%
    pull(sum_pred) %>% 
    append(pred_sums)
}

# actual probabilities
actual_sums <- list()
for (i in 1:20) {
  actual_sums <-
    hotel_fold_fit$.predictions[[i]] %>%
    summarize(sum_actual = sum(as.numeric(children))) %>%
    pull(sum_actual) %>% 
    append(actual_sums)
}
# colnames
names <- tibble("Folds" = c("Actual", "Predicted"))

probs <-
  as_tibble(actual_sums, .name_repair = "unique") %>% 
  rbind(as_tibble(pred_sums, .name_repair = "unique")) 
# Table
cbind(names, probs) %>%
  kable(col.names = 
          append("Folds", make.unique(c("Folds", rep("v", 21)), sep = "")[3:22]),
        caption = "Sum of probabilities")


tibble(fold = seq(1, 20, 1), 
       Actual = unlist(actual_sums), 
       Prediction = unlist(pred_sums)) %>%
  pivot_longer(!fold, names_to = "names", values_to = "vals") %>%
  ggplot() +
  geom_point(aes(x = fold, y = vals, color = names))  + 
  labs(x = "Fold", y = "Sum of probabilies") + 
  scale_x_continuous(breaks = seq(1, 20, 1)) +
  scale_color_brewer(palette = "Set1") + 
  theme_clean() + 
  theme(legend.title = element_blank())

mean_err <- sum(probs[1,] - probs[2,]) / 20
glue::glue("The average mean error is {mean_err}")
```
# The sum of probabilities is fairly similar across folds. 